\textbf{COMP0085 Summative Assignment}

### 2023-01-01 submission.tex
formatting
implement variational bayes for ard

We begin with the log joint:

\[\log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) = \log P(\textbf{x}|  \textbf{s}, \textbf{A}, \Psi,  \eta)
+ \log P( \textbf{s}| \pi, \eta) + \log P( \pi| \eta) + \log P( \textbf{A}| \eta) + \log P( \Psi| \eta)
\]

where $\textbf{A} \in \mathbb{R}^{D \times K}$ is a concatenation of $\mu_k$ for $k \in \{1, ..., K\}$  and \eta$ is a collection of all hyperparameters.

We know:

\[P(\textbf{x}|  \textbf{s}, \textbf{A}, \Psi,  \eta) = \frac{1}{(2\pi)^{\frac{D}{2}} |\Psi|^{\frac{1}{2}}} \exp\left(-\frac{1}{2}(\textbf{x}-\textbf{As})^T \Psi^{-1} (\textbf{x}-\textbf{As})\right)\]

\[P( \textbf{s}| \pi, \eta) = \prod_{k=1}^{K} \pi_k^{s_i} (1-\pi_k)^{1-s_k}\]

\[P( \pi| \eta) = \prod_{k=1}^{K} \frac{\pi_k^{\alpha - 1}(1-\pi_k)^{\beta-1}}{B(\alpha, \beta)}\]

For $\textbf{A}$ we choose a factorised conjugate prior for each column, $\mu_k$:

\[P( \textbf{A}| \eta) = \prod_{k=1}^{K}P( \mu_{k}| \eta)\]

For each column we choose:

\[P( \mu_{k}| \eta) = \mathcal{N}(\mu_{k}| \mu_{\mu_{k}}, \Sigma_{\mu_{k}}) =\frac{1}{(2\pi)^{\frac{D}{2}} |\Sigma_{\mu_{k}}|^{\frac{1}{2}}} \exp\left(-\frac{1}{2}(\mu_{k}} - \mu_{\mu_{k}})^T \Sigma_{\mu_{k}}^{-1}(\mu_{k} - \mu_{\mu_{k}})\right)\]

a Gaussian prior with diagonal covariance $\Sigma_{\mu_{k}} = \alpha_k^{2}\textbf{I}$ and mean zero, so we can simplify:

\[P( \mu_{k}| \alpha_k) = (2\pi\alpha_k^{2})^{\frac{-D}{2}} \exp\left(-\frac{\mu_{k}^T \mu_{k}}{2\alpha_k^{2}} \right)\]

For $\Psi}$ we choose a conjugate prior:

\[P( \Psi| \eta) = \prod_{d=1}^D InvGamma(\sigma^2| a, b) = \prod_{d=1}^D \frac{b^{a}}{\Gamma(a)}(\sigma^2)^{-a-1}\exp(-\frac{b}{\sigma^2})\]

a product of inverse gamma distributions, assuming $\Psi$ is a diagonal matrix with elements \sigma^2.

@@

\[
\begin{array}{l}
\log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) = \\
\tab \tab -\frac{-D}{2} \log (2 \pi \sigma^2)   -\frac{1}{2 \sigma^2} \left(\textbf{x}^T\textbf{x} - 2 \textbf{x}^T\sum_{k=1}^{K}  s_k  \mu_k   + \sum_{k=1}^{K} \sum_{k'=1, k'\neq k}^{K}  s_k s_{k'} \mu_k^T\mu_{k'} + \sum_{k=1}^{K}  s_k \mu_k^T\mu_k \right) \\
\tab \tab + \sum_{k=1}^{K} s_k \log \pi_k + (1-s_k) \log (1-\pi_k)}\\
\tab \tab + \sum_{i=1}^K (\alpha-1) \log \pi_k + (\beta - 1) \log (1-\pi_k) - \log B(\alpha, \beta)\\
\tab \tab + \sum_{i=1}^K -\frac{D}{2} \log (2 \pi\alpha_k^2)  -\frac{\mu_{k}^T \mu_{k}}{2\alpha_k^{2}}\\
\tab \tab + \sum_{d=1}^D a \log b + (-a - 1) \log(\sigma^2) - \frac{b}{\sigma^2} - \log \Gamma(a)\\
\end{array}
\]

%Simplifying:
%
%\[
%\begin{array}{l}
%\log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) = -\frac{D}{2} \log (2 \pi)  -\frac{1}{2}(\textbf{x}-\textbf{As})^T \Psi^{-1} (\textbf{x}-\textbf{As}) \\
%\tab \tab \tab \tab \tab + \sum_{i=1}^K (s_k+\alpha-1) \log \pi_k + (\beta - s_k) \log (1-\pi_k) - \log B(\alpha, \beta)\\
%\tab \tab \tab \tab \tab -\frac{DK}{2} \log (2 \pi) - \frac{1}{2} \log(|\Sigma_{\textbf{A}}|)   -\frac{1}{2}\textbf{A}^T \Sigma_{\textbf{A}}^{-1}\textbf{A}\\
%\tab \tab \tab \tab \tab + \sum_{d=1}^D a_d \log b_d + (-a_d - \frac{3}{2}) \log\sigma^2 - \frac{b_d}{\sigma^2} - \log \Gamma(a_d)\\
%\end{array}
%\]

@@

\[q_{\texbf{s}}(\textbf{s}) \propto \exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q(\theta)}\]

where $\theta$ denotes the parameters $\pi$, $\textbf{A}$, $\Psi$, $\eta$.

@@

\[q_{\texbf{s}}(\textbf{s}) \propto \exp \left\langle -\frac{1}{2 \sigma^2} \left( - 2 \textbf{x}^T\sum_{k=1}^{K}  s_k  \mu_k   + \sum_{k=1}^{K} \sum_{k'=1, k'\neq k}^{K}  s_k s_{k'} \mu_k^T\mu_{k'} + \sum_{k=1}^{K}  s_k \mu_k^T\mu_k \right) + \sum_{k=1}^{K} s_k \log \frac{\pi_k}{1-\pi_k}\right\rangle_{q(\theta)}\]

@@

\[\frac{\partial}{\partial q_i}\left( \exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q(\theta)} + \lambda^{LG} \int q_i -1)\right) \propto \exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q(\theta)\prod_{j\neq i} q_j(s_j)} - \log q_i(s_i)\]

@@

\[ \lambda_i =  \frac{1}{1+\exp\left[ - \left(\frac{  \langle \mu_i\rangle_{q_{\mu_i}}^T}{\langle\sigma^2\rangle_{q(\Psi)}} \left( \textbf{x} -  \frac{\langle \mu_i\rangle_{q_{\mu_i}}}{2} -  \sum_{j=1, j\neq i}^{K}   \lambda_j  \langle \mu_j\rangle_{q_{\mu_j} \right) + \log \frac{\langle \pi_i\rangle_{q_{\pi_i}}}{1-\langle \pi_i\rangle_{q_{\pi_i}}}\right) \right]}\]

@@

For the Variational Bayes maximisation step, we set:

\[q_{\theta}(\theta) \propto P(\theta) \exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q(\textbf{s})}\]

assuming the factorisation:

\[q_{\theta}(\theta) = q_{\pi}(\pi)q_{\Psi}(\Psi)q_{\textbf{A}}(\textbf{A})\]

we can calculate each factor independently.

For $q_{\pi}(\pi)$:

\[q_{\pi}(\pi) \propto P(\pi)\exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \pi}(\theta)}\]

@@

\[q_{\pi}(\pi) \propto \left(\prod_{k=1}^{K} \frac{\pi_k^{\alpha - 1}(1-\pi_k)^{\beta-1}}{B(\alpha, \beta)} \right)\exp \left\langle \sum_{k=1}^K s_k \log \pi_k + (1 - s_k) \log (1-\pi_k) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \pi}(\theta)}\]

We see:

\[q_{\pi}(\pi) \propto \prod_{k=1}^{K} \frac{\pi_k^{\alpha  + \frac{1}{D}\left(\sum_{k=1}^K \left\langle s_k\right\rangle_{q_{s_k}}\right) - 1}(1-\pi_k)^{\beta-\frac{1}{D}\left(\sum_{k=1}^K \left\langle s_k\right\rangle_{q_{s_k}}\right)}}{B(\alpha, \beta)}\]

\[q_{\pi}(\pi) = \prod_{k=1}^{K}  Beta\left(\alpha + \frac{1}{D}\sum_{k=1}^K \left\langle s_k\right\rangle_{q_{s_k}} , \beta+1- \frac{1}{D}\sum_{k=1}^K \left\langle s_k\right\rangle_{q_{s_k}}\right)\]

%Substituting the appropriate terms:
%
%\[q_{\pi}(\pi) \propto \left(\prod_{k=1}^{K} \frac{\pi_k^{\alpha - 1}(1-\pi_k)^{\beta-1}}{B(\alpha, \beta)} \right)\exp \left\langle \sum_{i=1}^K (s_k+\alpha-1) \log \pi_k + (\beta - s_k) \log (1-\pi_k) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \pi}(\theta)}\]
%
%We see:
%
%\[q_{\pi}(\pi) \propto \prod_{k=1}^{K} \frac{\pi_k^{2\alpha - 2 + \left\langle s_k\right\rangle_{q_{s_k}}}(1-\pi_k)^{2\beta-\left\langle s_k\right\rangle_{q_{s_k}}}}{B(\alpha, \beta)}\]
%
%\[q_{\pi}(\pi) = \prod_{k=1}^{K}  Beta(2\alpha -2 + \left\langle s_k\right\rangle_{q_{s_k}} , 2\beta+1-\left\langle s_k\right\rangle_{q_{s_k}}})\]

For $q_{\Psi}(\Psi)$:

\[q_{\Psi}(\Psi) \propto P(\Psi)\exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \Psi}(\theta)}\]

\[
q_{\Psi}(\Psi) \propto \left( \prod_{d=1}^D \frac{b^{a}}{\Gamma(a)}(\sigma^2)^{-a-1}\exp\left(-\frac{b}{\sigma^2}\right) \right)\exp \left\langle   -\frac{-D}{2} \log (\sigma^2)   -\frac{1}{2 \sigma^2} \left(\textbf{x}-\textbf{A}\textbf{s} \right)^T\left(\textbf{x}-\textbf{A}\textbf{s} \right)  \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \Psi(\theta)}}
\]

We see:

\[
q_{\Psi}(\Psi) \propto  \prod_{d=1}^D \frac{b^{a}}{\Gamma(a)}(\sigma^2)^{-(a+\frac{1}{2})-1}\exp\left(-\frac{b+\frac{2}{D} \left\langle \left(\textbf{x}-\textbf{A}\textbf{s} \right)^T\left(\textbf{x}-\textbf{A}\textbf{s} \right) \right\rangle_{q_{\textbf{s}}(\textbf{s})q_{\textbf{A}}(\textbf{A})}}{\sigma^2}\right) }
\]

where $\textbf{A}_{d :} \in \mathbb{R}^{1 \times K}$ is the $d^{th}$ row of $\textbf{A}$.

\[
q_{\Psi}(\Psi) = \prod_{d=1}^D  InvGamma\left(\sigma^2 \left| a+\frac{1}{2}, b+\frac{2}{D} \left\langle \left(\textbf{x}-\textbf{A}\textbf{s} \right)^T\left(\textbf{x}-\textbf{A}\textbf{s} \right) \right\rangle_{q_{\textbf{s}}(\textbf{s})q_{\textbf{A}}(\textbf{A})}\right)
\]

For $q_{\mu_{k}}(\mu_{k})$:

\[q_{\mu_{k}}(\mu_{k})  \propto P(\mu_{k})\exp \left\langle \log P(\textbf{x}, \textbf{s}, \pi, \textbf{A}, \Psi | \eta) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \mu_{k}}(\theta)}\]

\[q_{\mu_{k}}(\mu_{k}) \propto \exp\left(-\frac{\mu_{k}^T \mu_{k}}{2\alpha_k^{2}} \right) \exp \left\langle \frac{-1}{2\sigma^2} \left(- 2 \textbf{x}^T s_k  \mu_k  + s_k \mu_k^T\mu_k  + 2 \sum_{k'=1, k'\neq k}^{K}  s_k s_{k'} \mu_k^T\mu_{k'}}\right) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \mu_{k}}(\theta)}\]

\[q_{\mu_{k}}(\mu_{k}) \propto \exp\left(-\frac{1}{2}\left( \mu_{k}^T\left(\frac{1}{\alpha_k^{2}} + \frac{\langle s_k \rangle_{q_{s_k}}}{\langle\sigma^2\rangle_{q_{\Psi}}}}\right)\mu_{k} + 2 \langle s_k \rangle_{q_{s_k}}\left(-  \textbf{x}^T +  \sum_{k'=1, k'\neq k}^{K} \langle s_{k'} \rangle_{q_{s_{k'}}} \langle \mu_{k'} \rangle_{q_{\mu_{k'}}}^T\right)\mu_k\right)\]

We see that

\[q_{\mu_{k}}(\mu_{k}) = \mathcal{N}\left(\mu_{\mu_{k}}, \Sigma_{\mu_{k}}\right)}\]

@@

\[ \Sigma_{\mu_{k}} = \left(\frac{1}{\alpha_k^{2}} + \frac{\langle s_k \rangle_{q_{s_k}}}{\langle\sigma^2\rangle_{q_{\Psi}}}}\right)^{-1} \textbf{I}\]

@@

\[\mu_{\mu_{k}} = \Sigma_{\mu_{k}} \left(\textbf{x}^T - \sum_{k'=1, k'\neq k}^{K} \langle s_{k'} \rangle_{q_{s_{k'}}} \langle \mu_{k'} \rangle_{q_{\mu_{k'}}}^T\right)\langle s_k \rangle_{q_{s_k}}\]

Optimising the parameters $\{a, b\}$, and $\alpha^2$ in turn, will cause some $\alpha_i^2$ to diverge, the number of remaining $\alpha_i^2$ provide our determination for the value of $K$. This is automatic relevance determination through factor analysis.

@@

\newpage
\section*{Appendix 1: constants.py}
\lstinputlisting[language=Python]{src/constants.py}
\newpage
\section*{Appendix 2: main.py}
\lstinputlisting[language=Python]{main.py}
\end{document}

>>>>

We begin by writing the expression for $x_d$:

\[P(x_d | \texbf{s}, \textbf{w}_d, \sigma^2) = \mathcal{N}\left(\textbf{s}^T\textbf{w}_d, \sigma^2\right)\]

where we know from the diagonal covariance of $P(\textbf{x} | \textbf{s}, \mu, \sigma^2)$ that each dimension is independent. Moreover, $\textbf{w}_d \in \mathbb{R}^{K\times 1}$, which is the $d^{th}$ row of $\mu \in \mathbb{R}^{D \times K}$

Thus, we can write the posterior:

\[\log P(\textbf{x}, \textbf{s}, \mu| \pi , \sigma^2,  \alpha) = \log P( \textbf{s}| \pi) + \sum_{d=1}^{D} \log P(x_d | \texbf{s}, \textbf{w}_d, \sigma^2) + \log P( \textbf{w}_d| \alpha)
\]

where we introduce priors on each $\textbf{w}_k$ with $\alpha \in \mathbb{R}^{K \times 1}$.

We choose each prior to be:

\[P( \textbf{w}_d| \alpha) = \mathcal{N}(0, \textbf{A}^{-1})\]

where $\textbf{A} = diag(\alpha)$, the precision matrix.

@@

\[
\begin{array}{l}
\log P(\textbf{x}, \textbf{s}, \mu| \pi , \sigma^2,  \alpha) = \\
\tab \tab + \sum_{d=1}^D-\frac{-1}{2} \log (2 \pi \sigma^2)   -\frac{1}{2 \sigma^2} \left(x_d^2 - 2 x_d \textbf{s}^T \textbf{w}_d   + \textbf{w}_d^T \textbf{s}\textbf{s}^T\textbf{w}_d\right) \\
\tab \tab + \sum_{k=1}^{K} s_k \log \pi_k + (1-s_k) \log (1-\pi_k)}\\
\tab \tab + \sum_{d=1}^D -\frac{K}{2} \log (2 \pi) + \frac{1}{2}  \sum_{k=1}^K \left(\log\alpha_k \right)  -\frac{1}{2}\textbf{w}_{d}^T \textbf{A}\textbf{w}_{d}\\
\end{array}
\]

@@

\[q_{\texbf{s}}(\textbf{s}) \propto \exp \left\langle \log P(\textbf{x}, \textbf{s}, \mu| \pi , \sigma^2,  \alpha)  \right\rangle_{q(\mu)}\]

@@

\[q_{\texbf{s}}(\textbf{s}) \propto \exp \left\langle -\frac{1}{2 \sigma^2} \left( - 2 \textbf{x}^T\sum_{k=1}^{K}  s_k  \mu_k   + \sum_{k=1}^{K} \sum_{k'=1, k'\neq k}^{K}  s_k s_{k'} \mu_k^T\mu_{k'} + \sum_{k=1}^{K}  s_k \mu_k^T\mu_k \right) + \sum_{k=1}^{K} s_k \log \frac{\pi_k}{1-\pi_k}\right\rangle_{q(\mu)}\]

@@

\[\frac{\partial}{\partial q_i}\left( \exp \left\langle \log P(\textbf{x}, \textbf{s}, \mu| \pi , \sigma^2,  \alpha) \right\rangle_{q(\mu)} + \lambda^{LG} \int q_i -1)\right) \propto \exp \left\langle \log P(\textbf{x}, \textbf{s}, \mu| \pi , \sigma^2,  \alpha) \right\rangle_{q(\mu)\prod_{j\neq i} q_j(s_j)} - \log q_i(s_i)\]

@@

\[ \lambda_i =  \frac{1}{1+\exp\left[ - \left(\frac{  \langle \mu_i\rangle_{q_{\mu_i}}^T}{\sigma^2} \left( \textbf{x} -  \frac{\langle \mu_i\rangle_{q_{\mu_i}}}{2} -  \sum_{j=1, j\neq i}^{K}   \lambda_j  \langle \mu_j\rangle_{q_{\mu_j} \right) + \log \frac{ \pi_i}{1- \pi_i\right) \right]}\]

@@

For the maximisation step, we perform maximisation steps for the parameters $\sigma$ and $\pi$ in the same way as question 3. However, having defined a prior on $\mu$ (through $\textbf{w}$) so we will have to derive our expression for $\langle \mu_k\rangle_{q_{\mu_k}}$ the expectation of the posterior on $\mu_k$. This involves deriving the posterior distribution of $\textbf{w}_d$

\[q_{\textbf{w}_d}(\textbf{w}_d})  \propto P(\textbf{w}_d)\exp \left\langle \log P(\textbf{x}, \textbf{s}, \mu| \pi , \sigma^2,  \alpha) \right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \textbf{w}_d}(\textbf{w}_d)}\]

@@

\[q_{\textbf{w}_d}(\textbf{w}_d})\propto \exp\left(-\frac{1}{2}\textbf{w}_{d}^T \textbf{A}\textbf{w}_{d} \right) \exp \left\langle  -\frac{1}{2 \sigma^2} \left(- 2 x_d \textbf{s}^T \textbf{w}_d   + \textbf{w}_d^T \textbf{s}\textbf{s}^T\textbf{w}_d\right)\right\rangle_{q_{\textbf{s}}(\textbf{s}) q_{\neg \textbf{w}_d}(\textbf{w}_d)}\]

\[q_{\textbf{w}_d}(\textbf{w}_d})\propto \exp\left(-\frac{1}{2}\left(\textbf{w}_{d}^T \left(\textbf{A}+\frac{\langle  \textbf{s}\textbf{s}^T \rangle_{q_{\textbf{s}}(\textbf{s})}}{\sigma^2} \right)\textbf{w}_{d} - 2 \left(\frac{x_d \langle\textbf{s}^T \rangle_{q_{\textbf{s}}(\textbf{s})}}{\sigma^2}}\right) \textbf{w}_d  \right)\right)\right)\]

We see that the posterior:

\[q_{\textbf{w}_d}(\textbf{w}_d) = \mathcal{N}\left(\mu_{\textbf{w}_d}, \Sigma_{\textbf{w}_d}\right)}\]

@@

\[ \Sigma_{\textbf{w}_d} = \left(\frac{\langle  \textbf{s}\textbf{s}^T \rangle_{q_{\textbf{s}}(\textbf{s})}}{\sigma^2} + \textbf{A}\right)^{-1}\]

@@

\[\mu_{\textbf{w}_d} = \Sigma_{\textbf{w}_d}} \left(\frac{x_d \langle\textbf{s}^T \rangle_{q_{\textbf{s}}(\textbf{s})}}{\sigma^2}}\right)\]

Thus, $\langle \mu_k\rangle_{q_{\mu_k}} \in \mathbb{R}^{D \times 1}$ is simply the concatenation of the $k^{th}$ elements of $\mu_{\textbf{w}_d}$ for $d\in \{1,...,D\}$

For ARD, we must also optimise $\alpha$ with a hyper-M step. We start by choose $Ga(\alpha_k| a, b)$, a Gamma prior on \alpha_k$, with $a$ and $b$ being hyperparameters. Thus, to optimise $\alpha$ we want to maximise the penalised objective:

\[ \alpha =  \arg \max_{\alpha} \left\langle \log P(\textbf{x}, \textbf{s}, \mu| \pi , \sigma^2,  \alpha) \right\rangle_{q(\textbf{w})} + \sum_{k=1}^K \log P( \alpha_k| a, b)\]

Substituting the appropriate terms, we have our penalised objective $\mathcal{Q}$:

\[\mathcal{Q} =  \left\langle \sum_{d=1}^D \frac{1}{2}  \sum_{k=1}^K \left(\log\alpha_k \right)  -\frac{1}{2}\textbf{w}_{d}^T \textbf{A}\textbf{w}_{d} \right\rangle_{q(\textbf{w})} + \sum_{k=1}^K (a-1)\log\alpha_k -b\alpha_k\]

\[\mathcal{Q} =  \frac{D}{2}  \sum_{k=1}^K \left( \log\alpha_k \right)  - \frac{1}{2}\sum_{d=1}^D \left( tr \left[ \textbf{A}\left\langle  \textbf{w}_{d}\textbf{w}_{d}^T \right\rangle_{q(\textbf{w}_d)} \right]\right)+ \sum_{k=1}^K (a-1)\log\alpha_k -b\alpha_k\]

%Knowing $\left\langle  \textbf{w}_{d}\textbf{w}_{d}^T \right\rangle_{q(\textbf{w})} = \mu_{\textbf{w}_d}\mu_{\textbf{w}_d}^T + \Sigma_{\textbf{w}_d}$:
%
%\[\mathcal{Q} =  \frac{D}{2}  \sum_{k=1}^K \left( \log\alpha_k \right)  - \frac{1}{2}\sum_{d=1}^D \left( tr \left[ \textbf{A}\left(\mu_{\textbf{w}_d}\mu_{\textbf{w}_d}^T + \Sigma_{\textbf{w}_d}\right)} \right] \right)+ \sum_{k=1}^K (a-1)\log\alpha_k -b\alpha_k\]

Setting $\frac{d\mathcal{Q}}{d\alpha_k} = 0$ we get:

\[ \frac{D}{2\alpha_k} -\frac{1}{2} \sum_{d=1}^D\left\langle (w_{d, k})^2 \right\rangle_{q(\textbf{w}_{d})} + \frac{a-1}{\alpha_k} - b = 0\]

where $w_{d, k}$ is the $k^{th}$ element of $\textbf{w}_d$.

Knowing $\left\langle (w_{d, k})^2 \right\rangle_{q(\textbf{w}_{d})} = (\mu_{\textbf{w}_{d, k}})^2 + \Sigma_{\textbf{w}_{d, (k, k)}}$, we can solve for $\alpha_k$:

\[ \alpha_k = \frac{2a+D-2}{2b+ \sum_{d=1}^D \left((\mu_{\textbf{w}_{d, k}})^2 + \Sigma_{\textbf{w}_{d, (k, k)}}\right) }\]

we have our hyper-M steps for optimising $\alpha$.

%Optimising the parameters $\{a, b\}$, and $\alpha^2$ in turn, will cause some $\alpha_i^2$ to diverge, the number of remaining $\alpha_i^2$ provide our determination for the value of $K$. This is automatic relevance determination through factor analysis.

\newpage
The Python code for variational Bayes:
\lstinputlisting[language=Python]{src/models/variational_bayes.py}

\newpage
The rest of the Python code for question 4:
\lstinputlisting[language=Python]{src/solutions/q4.py}

@@

\newpage
\section*{Appendix 1: constants.py}
\lstinputlisting[language=Python]{src/constants.py}
\newpage
\section*{Appendix 2: main.py}
\lstinputlisting[language=Python]{main.py}
\newpage
\section*{Appendix 3: generate\_images.py}
\lstinputlisting[language=Python]{src/generate_images.py}
\newpage
\section*{Appendix 4: MStep.py}
\lstinputlisting[language=Python]{demo_code/MStep.py}
\end{document}
